---
title: "ROSMAP_MT"
author: "Raymond Yuan SHANG"
date: "8/1/2019"
output: html_document
---
```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages(c("dynamicTreeCut", "cluster", "flashClust", "Hmisc", "reshape", "foreach", "doParallel") ) 
#source("http://bioconductor.org/biocLite.R") 
#biocLite("WGCNA")
library(WGCNA)
library(cluster)
library(flashClust)
library("readxl")
library("openxlsx")
library(clusterProfiler)
library(org.Hs.eg.db)
library(org.Rn.eg.db)
library(ggplot2)
library(venn)
library(tidyverse)
library(limma)
library("edgeR")
library(org.Hs.eg.db)

options(stringsAsFactors = FALSE)
#theme_set(theme_bw())
theme_set(theme_bw()) + theme_update(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),plot.title = element_text(hjust = 0.5),legend.title=element_blank()) #legend.position = c(0.8, 0.8),
```

```{r}
rm(list=ls())
keytypes(org.Hs.eg.db)
```

```{r}
wrapper <- function(x, ...) 
{
  paste(strwrap(x, ...), collapse = "\n")
}

ggplotRegression <- function (fit) {

require(ggplot2)

ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  labs(title=fit$genename,subtitle = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, digits=5),
                     "Intercept =",signif(fit$coef[[1]],5 ),"\n",
                     "Slope =",signif(fit$coef[[2]], 5),
                     " P =",signif(summary(fit)$coef[2,4], 5)))
}
plot(density(rnorm(10000)))
```



##Define geneSets
```{r}
###human
complex1_human <- c( "MT-ND1","MT-ND2","MT-ND3","MT-ND4","MT-ND4L","MT-ND5","MT-ND6",
                      "NDUFS1","NDUFS2","NDUFS3","NDUFS7","NDUFS8","NDUFV1","NDUFV2","NDUFAB1","NDUFA1","NDUFA2","NDUFA3","NDUFA5","NDUFA6",
                      "NDUFA7","NDUFA8","NDUFA9","NDUFA10","NDUFA11","NDUFA12","NDUFA13","NDUFB1","NDUFB2","NDUFB3","NDUFB4",
                      "NDUFB5","NDUFB6","NDUFB7","NDUFB8","NDUFB9","NDUFB10","NDUFB11","NDUFC1","NDUFC2","NDUFS4","NDUFS5","NDUFS6","NDUFV3")

complex2_human <- c("SDHA","SDHB","SDHC","SDHD")


complex3_human <- c("MT-CYB","CYC1","UQCRB","UQCRC1","UQCRC2","UQCRFS1","UQCRH","UQCRQ","UQCR10","UQCR11")


complex4_human <- c("MT-CO1","MT-CO2","MT-CO3",
                    "COX4I1","COX4I2","COX5A","COX5B","COX6A1","COX6A2","COX6B1","COX6B2","COX6C",
                    "COX7A1","COX7A2","COX7B","COX7B2","COX7C","COX8A","COX8C")

complex5_human <-c("MT-ATP6","MT-ATP8",
                   "ATP5F1A","ATP5F1B","ATP5F1C","ATP5F1D","ATP5F1E","ATP5MC1","ATP5MC2","ATP5MC3",
                    "ATP5MD","ATP5ME","ATP5MF","ATP5MG","ATP5MPL","ATP5PB","ATP5PD","ATP5PF","ATP5PO","ATP5IF1")

MT_human <- c("MT-ND1","MT-ND2","MT-ND3","MT-ND4","MT-ND4L","MT-ND5","MT-ND6","MT-CYB","MT-CO1","MT-CO2","MT-CO3","MT-ATP6","MT-ATP8")

synapseMarkers <- c("RAB3C","RAB27B","SYNJ1","GAD2","CALB1","DNAJC6","DCLK1","GABRG2")
#others_human <- c("Ndufa4","Ndufa4l2","Ndufaf1","Ndufaf2","Ndufaf3","Ndufaf4")

```

```{r}
##BrainCellBioMarker_human.xlsx
#BrainNeuronStemCellBioMarker_human.xlsx
cellMarkers <- read_excel("~/Dropbox/RaymondTools/others/BrainCellBioMarker_human.xlsx",sheet=1) #BrainNeuronStemCellBioMarker_human.xlsx
colnames(cellMarkers)
# [1] "Neuroepithelial"              "SchwannCellPrecursor"         "MyelinatingSchwannCell"       "NonMyelinatingSchwannCell"    "RadialGlia"                   "OligodendrocytePrecursorCell" "Oligodendrocytes"            
# [8] "Astrocytes"                   "Microglia"                    "IntermediateProgenitors"      "ImmatureNeurons"              "MatureNeurons"                "GlutamatergicNeurons"         "GABAergic Neurons"           
#[15] "Dopaminergic Neurons"         "Serotonergic Neurons"         "Cholinergic Neurons" 
cellMarkers <- as.data.frame(cellMarkers)
celltypes <- colnames(cellMarkers)
celltypes
```


##read demographic information
```{r}
s2c4 <- read.csv(file = "~/Dropbox/human/ROSMAP/RNASeq/demographic_tidy2.csv",stringsAsFactors = FALSE)
head(s2c4$rnaseq_id)
colnames(s2c4)

t2g_all <- read.csv(file="~/Dropbox/RaymondTools/DEG/tx2gene_genecodeV24_humanV3.txt",header=TRUE,stringsAsFactors = FALSE)
head(t2g_all)
t2g <- t2g_all[,c(1,2)]
head(t2g)
```







#ROSMAP
```{r}
#ROSMAP_DLPFC_netResidualExpression.tsv
plotDistribution = function (x) {
  N = length(x)
  x <- na.omit(x)
  hist( x,col = "light blue",
        probability = TRUE)
  lines(density(x), col = "red", lwd = 3)
  rug(x)
  print(N-length(x))
}

#MAYO_CBE_TCX_netResidualExpression.tsv
#MSSM_FP_STG_PHG_IFG_netResidualExpression.tsv
#ROSMAP_DLPFC_netResidualExpression.tsv

dir_ROSMAPReseq="~/Dropbox/human/RNAResequence/ROSMAP"  #MayoRNAseq,MSBB,ROSMAP

```



##read row counts data
```{r}
#ROSMAP_DLPFC_eCounts.tsv,ROSMAP_DLPFC_Counts.tsv
expr_ROSMAP_counts <- read.delim2(file=paste0(dir_ROSMAPReseq,"/","ROSMAP_DLPFC_Counts.tsv"))
str(expr_ROSMAP_counts[1:5,1:5])
rownames(expr_ROSMAP_counts) <- expr_ROSMAP_counts[,1]
expr_ROSMAP_counts <- expr_ROSMAP_counts[,-1]
lapply(expr_ROSMAP_counts,as.numeric,2) -> expr_ROSMAP_counts2
data.frame(expr_ROSMAP_counts2,check.names = F) -> expr_ROSMAP_counts2
rownames(expr_ROSMAP_counts2) <- rownames(expr_ROSMAP_counts)

expr_ROSMAP_counts2[1:5,1:5]
```



#make DGEList object, clean data
```{r}
cols_filter <- colnames(expr_ROSMAP_counts2)    #632 samples
cols_filter[cols_filter %in% paste0("X",s2c4$rnaseq_id)] ->cols_filter2  ##614 samples
s2c4 -> s2c4_ROSMAP_yshang
rownames(s2c4_ROSMAP_yshang) <- paste0("X",s2c4_ROSMAP_yshang$rnaseq_id)
s2c4_ROSMAP_yshang[cols_filter2,] ->s2c4_ROSMAP_yshang


x <- DGEList(expr_ROSMAP_counts2[,cols_filter2])
x[1:5,1:5]
cols_x <- colnames(x)
cols_x == rownames(s2c4_ROSMAP_yshang)
x$samples$group <- s2c4_ROSMAP_yshang$AD

#Batch,apoe_genotype, msex, AD
x$samples$Batch <- factor(s2c4_ROSMAP_yshang$Batch)
x$samples$apoe_genotype <- factor(s2c4_ROSMAP_yshang$apoe_genotype)
x$samples$msex <- factor(s2c4_ROSMAP_yshang$msex)
x$samples$pmi <- as.numeric(s2c4_ROSMAP_yshang$pmi)
x$samples$AOD <- as.numeric(s2c4_ROSMAP_yshang$age_death)
x$samples$RIN <- as.numeric(s2c4_ROSMAP_yshang$RINcontinuous)
x$samples$RIN2 <- as.numeric(s2c4_ROSMAP_yshang$RINcontinuous * s2c4_ROSMAP_yshang$RINcontinuous)
x$samples$groupCom <- paste0(x$samples$group,".",x$samples$msex,".",x$samples$apoe_genotype)
x$samples$AOD2 <- cut(x$samples[,"AOD"], breaks=c(0, 75,80,85,90,Inf),right = FALSE)
x$samples$AOD2 <- gsub('\\[',"",as.character(x$samples$AOD2))
x$samples$AOD2 <- gsub('\\)',"",as.character(x$samples$AOD2))
x$samples$AOD2 <- gsub('\\,',"_",as.character(x$samples$AOD2))
x$samples$Braak <- as.numeric(s2c4_ROSMAP_yshang$braaksc)
x$samples$rna_id <- s2c4_ROSMAP_yshang$rnaseq_id
class(x)
dim(x)
#x_pre <- x
x <- x[,x$samples$apoe_genotype %in% c("APOE33","APOE23","APOE34") ]
x$samples$apoe_genotype <- droplevels(x$samples$apoe_genotype)
dim(x)

genes <- as.data.frame(rownames(x$counts),stringsAsFactors=FALSE)
colnames(genes) <- "ENSEMBL"  #"HGNC_SYMBOL"
genes$ENSEMBL[!(genes$ENSEMBL %in% t2g_all$GENEID)]
#"ENSG00000273748" "ENSG00000271254" "ENSG00000278817" "ENSG00000277196"


id_gene <- unique(t2g_all[,c(3,2,4,5,6)])
id_gene <- id_gene[!duplicated(id_gene$GENEID),]
id_gene$ENSEMBL <- id_gene$GENEID
#tmp_merge <- merge(genes,id_symbol,by.x="HGNC_SYMBOL",by.y='HGNC_SYMBOL',all.x=TRUE)
genes_df <- dplyr::left_join(genes,id_gene,by="ENSEMBL")
x$genes <- dplyr::slice(genes_df,match(genes$ENSEMBL,genes_df$ENSEMBL))

length(x$genes$ENSEMBL ) - length(unique(x$genes$ENSEMBL)) ##Should be 0, otherwise, not 1:1 
dim(x)  # 15582 588

"COX1" %in% x$genes$HGNC_SYMBOL
"ENSG00000131055" %in% x$genes$ENSEMBL
complex1_human[!(complex1_human %in% x$genes$HGNC_SYMBOL)]
#"ENSG00000178127","ENSG00000267855"
#"NDUFV2","NDUFA7"

complex2_human[!(complex2_human %in% x$genes$HGNC_SYMBOL)]
complex3_human[!(complex3_human %in% x$genes$HGNC_SYMBOL)]
complex4_human[!(complex4_human %in% x$genes$HGNC_SYMBOL)]
#"ENSG00000131055",
#"COX4I2" "COX6A2" "COX6B2" "COX7B2" "COX8C" 
complex5_human[!(complex5_human %in% x$genes$HGNC_SYMBOL)]
dim(x)   #15582 588
```

##Before normalization
```{r}
library(RColorBrewer)
y=x
col <- brewer.pal( dim(x)[2], "Paired")
lcpm_y <- cpm(y, log=TRUE,prior.count = 0.25)  #0.25
{
png("Fig1/BeforeNormalization Lcpm.png",width = 1280, height = 960)
plot(density(lcpm_y[,1]), col=col[1], lwd=2, ylim=c(0,0.21), las=2, 
     main="", xlab="")
title(main=paste0("Before Normalization (first 12 samples:",length(lcpm_y)/588,")genes"), xlab="Log-cpm")
abline(v=0, lty=3)
for (i in 1:12){
   den <- density(lcpm_y[,i])
   lines(den$x, den$y, col=col[i], lwd=2)
}
dev.off()
}
```

##After normalization
```{r}
x <- calcNormFactors(x, method = "TMM")
x$samples$norm.factors
{
  png("Fig1/Normalization_lcpm.png",width = 1280, height = 600)
par(mfrow=c(1,2))
{
  plot(density(lcpm_y[,1]), col=col[1], lwd=2, ylim=c(0,0.21), las=2, 
     main="", xlab="")
title(main=paste0("Before Normalization (first 12 samples:",length(lcpm_y)/588,")genes"), xlab="Log-cpm")
abline(v=0, lty=3)
for (i in 1:12){
   den <- density(lcpm_y[,i])
   lines(den$x, den$y, col=col[i], lwd=2)
}
}
{
lcpm_x <- cpm(x, log=TRUE,prior.count = 0.25)  #5
plot(density(lcpm_x[,1]), col=col[1], lwd=2, ylim=c(0,0.21), las=2, 
     main="", xlab="")
title(main=paste0("After Normalization:first 12 samples(",length(lcpm_x)/588,") genes"), xlab="Log-cpm")
abline(v=0, lty=3)
for (i in 2:12){
   den <- density(lcpm_x[,i])
   lines(den$x, den$y, col=col[i], lwd=2)
}


}

dev.off()
}
```

##save results
```{r}
x_ROSMAP =x
lcpm_x_ROSMAP = lcpm_x
```

## Remove covariates
###Plot lcpm ~ covariante to check potential effects
```{r}
y=x_ROSMAP
lcpm_mds <- lcpm_x_ROSMAP
MT_genes <- MT_human
MT_genes_ENSG <- dplyr::filter(x_ROSMAP$genes,x_ROSMAP$genes$HGNC_SYMBOL %in% MT_human)
MT_genes_ENSG <- MT_genes_ENSG$ENSEMBL
MT_genes_ENSG %>% length()

expr_currentGene <- lcpm_mds["ENSG00000198888",]  #ENSG00000198888, MT-ND1, c("ENSG00000198888","ENSG00000198763")
expr_currentGene2 <- lcpm_mds["ENSG00000198727",]  #ENSG00000198727 MT-CYB   #ENSG00000125356 NDUFA1 #ENSG00000198763 ND2
expr_currentGene3 <- lcpm_mds["ENSG00000125356",]  #NDUFA1
plot(density(t(expr_currentGene)))

 df_current <- data.frame(sample=names(expr_currentGene),lcpm = t(t(expr_currentGene)) )
 df_current$lcpm <- t(t(expr_currentGene))
 df_current$gene2 <- t(t(expr_currentGene2))
  df_current$gene3 <- t(t(expr_currentGene3))

 df_current$AOD <- x_ROSMAP$samples$AOD[match(df_current$sample,rownames(x_ROSMAP$samples))]
df_current$AOD2 <- x_ROSMAP$samples$AOD2[match(df_current$sample,rownames(x_ROSMAP$samples))]

 df_current$AD <- x_ROSMAP$samples$groupCom[match(df_current$sample,rownames(x_ROSMAP$samples))]
 df_current$AD <- gsub("\\..*","",df_current$AD)
 df_current$RIN <- x_ROSMAP$samples$RIN[match(df_current$sample,rownames(x_ROSMAP$samples))]
df_current$PMI <- x_ROSMAP$samples$pmi[match(df_current$sample,rownames(x_ROSMAP$samples))]
df_current$Sex <- x_ROSMAP$samples$msex[match(df_current$sample,rownames(x_ROSMAP$samples))]
df_current$Braak <- x_ROSMAP$samples$Braak[match(df_current$sample,rownames(x_ROSMAP$samples))]
df_current$APOE <- x_ROSMAP$samples$apoe_genotype[match(df_current$sample,rownames(x_ROSMAP$samples))]
df_current$Batch <- x_ROSMAP$samples$Batch[match(df_current$sample,rownames(x_ROSMAP$samples))]

 plot(lcpm ~ AOD, df_current,main="Age of death")
 plot(lcpm ~ RIN, df_current,main="RIN")
 plot(lcpm ~ Batch, df_current,main="Batch")
 plot(lcpm ~ Sex, df_current,main="Sex")
 plot(lcpm ~ APOE, df_current,main="APOE")
 plot(lcpm ~ Braak, df_current,main="Braak")
 plot(lcpm ~ PMI, df_current,main="PMI")
 
 
 plot(lcpm ~ gene2, df_current)
 plot(lcpm ~ gene3, df_current)


 ggplot(data=df_current[df_current$Sex == "M",],aes_string(x="AOD",y="lcpm",col="AD"))+
   geom_point()
 
  p<- ggplot(data=df_current,aes_string(x="RIN",y="lcpm",col="Batch"))+
   geom_point(size=4)
  ggsave(p,filename = "Fig1/BatchRIN.png")
  
  ggplot(data=df_current,aes_string(x="RIN",y="lcpm",col="Sex"))+
   geom_point(size=4)
  
   ggplot(data=df_current,aes_string(x="RIN",y="lcpm",col="AD"))+
   geom_point(size=4)
   
   ggplot(data=df_current,aes_string(x="RIN",y="lcpm",col="Braak"))+
   geom_point(size=4)
   
   ggplot(data=df_current,aes_string(x="RIN",y="lcpm",col="PMI"))+
   geom_point(size=4)
   
 
```

### Regress out effects from coviants

```{r}

#Test residues
lm_fit <- lm(lcpm ~ RIN+Batch, data=df_current)
df_current$residual <- resid(lm_fit)
df_current$fitted <- fitted(lm_fit)


summary(lm_fit)

{
  pdf("Fig1/residual_test.pdf",width=12,height=12)
plot(resid(lm_fit) ~ fitted(lm_fit),#
main="Residual plot (ND1,Residual to Fitted,RIN+Batch)")
abline(h=0)
dev.off()
}

  ggplot(data=df_current,aes_string(x="fitted",y="residual",col="AD"))+
   geom_point(size=4)
  {
    plot(density(df_current$fitted[df_current$AD=="NCI"]),col="green")
    lines(density(df_current$fitted[df_current$AD=="MCI"]),col="blue")
    lines(density(df_current$fitted[df_current$AD=="AD"]),col="red")
   }
  
  
  {
    plot(density(df_current$fitted[df_current$Braak %in% c(1,2) & df_current$AD == "AD" ]) ,col="green")
    lines(density(df_current$fitted[df_current$Braak %in% c(3,4) & df_current$AD == "AD"] ) ,col="blue")
    lines(density(df_current$fitted[df_current$Braak %in% c(5,6) & df_current$AD == "AD"] ) ,col="red")
  }
  
  {
    plot(density(df_current$fitted[df_current$AD=="NCI" &  df_current$Braak %in% c(5,6)]),col="green")
    lines(density(df_current$fitted[df_current$AD=="MCI" & df_current$Braak %in% c(5,6)]),col="blue")
    lines(density(df_current$fitted[df_current$AD=="AD" & df_current$Braak %in% c(5,6)]),col="red")
  }
  
```




```{r}
df_current$Braak <- as.factor(df_current$Braak)
  ggplot(data=df_current[df_current$RIN>=6.5 &df_current$AD=="AD" ,],aes_string(x="Braak",y="lcpm",col="AD"))+ #AD
   geom_point()
{
  p1 <- ggplot(data=df_current[df_current$RIN>=0.5 &df_current$AD=="AD" ,],aes_string(x="Braak",y="lcpm",col="Sex"))+ #AD
          ylim(11,16)+ #geom_point()+
          #geom_boxplot(outlier.shape = NA) +
          #geom_jitter(binaxis='y', stackdir='center',position=position_jitter(0.1),size=3)
          geom_boxplot(aes(color=Sex), width=0.5,size=0.4,position=position_dodge((0.8)))+
          geom_dotplot(aes(fill=Sex,color=Sex), trim=F,binaxis = "y",stackdir = "center",dotsize = 0.4,position = position_dodge(0.8))
  p1
  
  p2 <- ggplot(data=df_current[df_current$RIN>=0.5 &df_current$AD=="MCI",],aes_string(x="Braak",y="lcpm",col="Sex"))+ #AD
          ylim(11,16)+
          geom_boxplot(aes(color=AD), width=0.5,size=0.4,position=position_dodge((0.8)))+
          geom_dotplot(aes(fill=Sex,color=Sex), trim=F,binaxis = "y",stackdir = "center",dotsize = 0.4,position = position_dodge(0.8))
  
  
  p3 <- ggplot(data=df_current[df_current$RIN>=0.5 &df_current$AD=="NCI",],aes_string(x="Braak",y="lcpm",col="Sex"))+ #AD
          ylim(11,16)+
          geom_boxplot(aes(color=Sex), width=0.5,size=0.4,position=position_dodge((0.8)))+
          geom_dotplot(aes(fill=Sex,color=Sex), trim=F,binaxis = "y",stackdir = "center",dotsize = 0.4,position = position_dodge(0.8))
  
   }
#library(cowplot)
#install.packages("ggpubr")
#library(ggpubr)
  library(tidyverse)
  {
    #pdf("CYB_SexDifferenceBasedOnBraak.pdf")
    ggarrange(p3, p2, p1 + rremove("x.text"), 
          labels = c("NCI", "MCI", "AD"),
          ncol =3, nrow = 1) %>% ggexport(filename = "ND1_SexDifferenceBasedOnBraak.png",width = 1440, height = 480)
    #dev.off()
  }
  
  sum(df_current$RIN>=6.5 & df_current$Sex=="M" & df_current$AD =="AD" & df_current$APOE =="APOE33")
```



```{r,eval=TRUE}
y=x_ROSMAP
lcpm_mds <- lcpm_x_ROSMAP
covvarianceM <- model.matrix(~y$samples$pmi+y$samples$Batch+y$samples$RIN+y$samples$AOD2) #
design_withoutFactors <- model.matrix(~y$samples$groupCom)


colnames(design_withoutFactors) <- gsub("y\\$samples\\$", "", colnames(design_withoutFactors))
head(design_withoutFactors)
colnames(design_withoutFactors) <- gsub("apoe_genotype", "", colnames(design_withoutFactors))
head(design_withoutFactors)
colnames(design_withoutFactors) <- gsub("group", "", colnames(design_withoutFactors))
head(design_withoutFactors)
rownames(design_withoutFactors) <- rownames(y$samples)
head(design_withoutFactors)
lcpm_WithoutFactors <- removeBatchEffect(lcpm_mds,covariates=covvarianceM,design=design_withoutFactors) #lcpm_mds
#save(lcpm_WithoutFactors,file="~/Dropbox/human/ROSMAP/RNASeq/lcpm_WithoutFactors_ROSMAP.Rdata")
#save(y,file="~/Dropbox/human/ROSMAP/RNASeq/x_normalized_ROSMAP.Rdata")
#Partial NA coefficients for 23365 probe(s)
#genes_detected <- rownames(lcpm_WithoutFactors)
for(i in 1:length(rownames(lcpm_WithoutFactors))){
  if(sum(is.na(lcpm_WithoutFactors[i,])) > 0){
    print(rownames(lcpm_WithoutFactors) [i])
  }
}


if(FALSE){
samples_tmp <- merge(y$samples,s2c4_ROSMAP_yshang,by.x="rna_id",by.y="rnaseq_id",all.x=TRUE)
colnames(samples_tmp)
samples_tmp <- samples_tmp[,c(1:15,21)]
samples_tmp <- samples_tmp[match(y$samples$rna_id,samples_tmp$rna_id),]
y$samples <- samples_tmp
}
```


###Plot (fitted + residuals) with the variable again
```{r}
{
  df_rest <- data.frame(sample = names(lcpm_WithoutFactors["ENSG00000198888",]),
                        gene1 = lcpm_WithoutFactors["ENSG00000198888",],  #MT-ND1
                        gene2 = lcpm_WithoutFactors["ENSG00000198727",],  #MT-CYB
                        gene3 = lcpm_WithoutFactors["ENSG00000132535",]    
  )
  
  df_rest$AOD <- x_ROSMAP$samples$AOD[match(df_rest$sample,rownames(x_ROSMAP$samples))]
df_rest$AOD2 <- x_ROSMAP$samples$AOD2[match(df_rest$sample,rownames(x_ROSMAP$samples))]

df_rest$AD <- x_ROSMAP$samples$groupCom[match(df_rest$sample,rownames(x_ROSMAP$samples))]
df_rest$AD <- gsub("\\..*","",df_rest$AD)
df_rest$RIN <- x_ROSMAP$samples$RIN[match(df_rest$sample,rownames(x_ROSMAP$samples))]
df_rest$PMI <- x_ROSMAP$samples$pmi[match(df_rest$sample,rownames(x_ROSMAP$samples))]
df_rest$Sex <- x_ROSMAP$samples$msex[match(df_rest$sample,rownames(x_ROSMAP$samples))]
df_rest$Braak <- x_ROSMAP$samples$Braak[match(df_rest$sample,rownames(x_ROSMAP$samples))]
df_rest$APOE <- x_ROSMAP$samples$apoe_genotype[match(df_rest$sample,rownames(x_ROSMAP$samples))]
df_rest$Batch <- x_ROSMAP$samples$Batch[match(df_rest$sample,rownames(x_ROSMAP$samples))]
  
##Plot Fitted data with RIN and Batch
{
   p<- ggplot(data=df_rest,aes_string(x="RIN",y="gene1",col="Batch"))+
      geom_point(size=4)+ylab("MT-ND1")
    ggsave(p,filename = "Fig1/BatchRIN_afterVarRM.png",width = 7, height = 7)
}

##Plot MT-ND1 and MT-ND2
lm_fit <- lm(gene1 ~ gene2, data=df_rest)
p <- ggplotRegression(lm_fit) +
                 xlab("MT-CYB") +
                ylab("MT-ND1")
p

ggsave(p,filename = "Fig1/ND1 and CYB correlations(p < 2E-284).png",width = 3, height =3)
    
    
##Plot MT-ND1 and NDUFA1
lm_fit <- lm(gene1 ~ gene3, data=df_rest[df_rest$AD == "NCI",])
p <- ggplotRegression(lm_fit) +
                 xlab("DLG4") +
                ylab("MT-ND1")
p
ggplot(data=df_rest,aes_string(x="gene3",y="gene1",col="Sex")) +
  geom_point()
ggsave(p,filename = "Fig1/ND1 and NDUFA1 correlations.png",width = 3, height =3)

##Plot MT-ND1 with AD
df_rest$AD <- factor(df_rest$AD,levels =c("NCI","MCI","AD"))
if(FALSE){
    png("Fig1/MT-ND1 density map after regression.png",width=1280,height=1024)
    plot(density(df_rest$gene1[df_rest$AD=="NCI"]),col="green",main="MT-ND1")
    #lines(density(df_rest$gene1[df_rest$AD=="NCI"]),col="green")
    lines(density(df_rest$gene1[df_rest$AD=="MCI"]),col="blue")
    lines(density(df_rest$gene1[df_rest$AD=="AD"]),col="red")
    dev.off()
  }

  p <- ggplot(df_rest,aes(x=gene1,col=AD))  +scale_colour_manual(values = c("green", "blue", "red"))+
    xlim(18,22)+xlab("lcpm (MT-ND1)") + geom_line(stat="density")+ggtitle("MT-ND1 expression density")+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position = c(0.8, 0.8),plot.title = element_text(hjust = 0.5),legend.title=element_blank())
  ggsave(p,filename = "Fig1/MT-ND1 density map after regression_AllInONE_AD.png",width = 5, height =3)
  #+ labs(x = NULL) +
     #opts(legend.position = "none") + opts(title = "Densities from a kernel density estimator")
  
  
  
  ##Plot MT-ND1 with Braak stage
  df_rest$Braak2 <- df_rest$Braak
  
  df_rest$Braak2[df_rest$Braak2 ==0 | df_rest$Braak2 ==1]="low"
  df_rest$Braak2[df_rest$Braak2 ==2 | df_rest$Braak2 ==3 |df_rest$Braak2 ==4]="middle"
  df_rest$Braak2[df_rest$Braak2 ==5 | df_rest$Braak2 ==6]="high"
df_rest$Braak2 <- factor(df_rest$Braak2,levels=c("low","middle","high"))
  p <- ggplot(df_rest,aes(x=gene1,col=Braak2))  +scale_colour_manual(values = c("green", "blue", "red"))+
    xlim(18,22)+xlab("MT-ND1") + geom_line(stat="density")+ggtitle("MT-ND1 expression density vs BraakStage")+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position = c(0.8, 0.8),plot.title = element_text(hjust = 0.5),legend.title=element_blank())
  ggsave(p,filename = "Fig1/MT-ND1 density map after regression_AllInONE_Braak.png",width = 5, height =3)

    ##Plot MT-ND1 with Sex
  p <- ggplot(df_rest,aes(x=gene1,col=Sex))  +scale_colour_manual(values = c("red", "green"))+
    xlim(18,22)+xlab("MT-ND1") + geom_line(stat="density")+ggtitle("MT-ND1 expression density vs Sex")+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position = c(0.8, 0.8),plot.title = element_text(hjust = 0.5),legend.title=element_blank())
  ggsave(p,filename = "Fig1/MT-ND1 density map after regression_AllInONE_Sex.png",width = 5, height =3)
  
      ##Plot MT-ND1 with APOE
  p <- ggplot(df_rest,aes(x=gene1,col=APOE))  +scale_colour_manual(values = c("red","blue", "green"))+
    xlim(18,22)+xlab("MT-ND1") + geom_line(stat="density")+ggtitle("MT-ND1 expression density vs Sex")+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position = c(0.8, 0.8),plot.title = element_text(hjust = 0.5),legend.title=element_blank())
  ggsave(p,filename = "Fig1/MT-ND1 density map after regression_AllInONE_APOE.png",width = 5, height =3)
  
  ##Plot MT-ND1 with Sex.AD
  df_rest$SexAD = paste0(df_rest$Sex,".",df_rest$AD)
    p <- ggplot(df_rest,aes(x=gene1,col=SexAD))  + #scale_colour_manual(values = c("red","blue", "green"))+
    xlim(18,22)+xlab("MT-ND1") + geom_line(stat="density")+ggtitle("MT-ND1 expression density vs Sex")+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position = c(0.8, 0.7),plot.title = element_text(hjust = 0.5),legend.title=element_blank())
  ggsave(p,filename = "Fig1/MT-ND1 density map after regression_AllInONE_SexAD.png",width = 5, height =4)
  
  ##Plot Percentage change of NCI, MCI and AD vs MT-ND1
  prior_AD <- sum(df_rest$AD == "AD")/length(df_rest$AD)
  prior_MCI <- sum(df_rest$AD == "MCI")/length(df_rest$AD)
  prior_NCI <- sum(df_rest$AD == "NCI")/length(df_rest$AD)
  tmp_AD <- density(df_rest$gene1[df_rest$AD == "AD"])
  tmp_NCI <- density(df_rest$gene1[df_rest$AD == "NCI"])
  df_rest_tmp <- df_rest
  
  df_rest_tmp[order(df_rest_tmp$gene1),] -> df_rest_tmp
  bin_size=100;lcpm_gene1=c();ratio_AD=c();ratio_MCI=c();ratio_NCI=c();
  lcpm_set <- c()
  current_count=0
  current_AD=current_MCI=current_NCI=0
  
  for(i in 1:length(df_rest_tmp$sample)){
    current_count <- current_count+1
    lcpm_set <- c(lcpm_set, df_rest_tmp$gene1[i])
    if(df_rest_tmp$AD[i] == "AD"){
      current_AD <- current_AD +1
    }else if(df_rest_tmp$AD[i] == "MCI"){
      current_MCI <- current_MCI +1
    }else{
      current_NCI <- current_NCI+1
    }
    if(current_count ==bin_size){
      lcpm_gene1 <- c(lcpm_gene1, mean(lcpm_set))
      ratio_AD <- c(ratio_AD,current_AD/bin_size/prior_AD)
      ratio_MCI <- c(ratio_MCI,current_MCI/bin_size/prior_MCI)
      ratio_NCI <- c(ratio_NCI,current_NCI/bin_size/prior_NCI)
      current_count=0;current_NCI=current_MCI=current_AD=0;
      lcpm_set <- c();
    }
  }
  ####After the loop, conclude the remaining
  {
    length(lcpm_set)
    lcpm_gene1 <- c(lcpm_gene1, mean(lcpm_set))
      ratio_AD <- c(ratio_AD,current_AD/bin_size/prior_AD)
      ratio_MCI <- c(ratio_MCI,current_MCI/bin_size/prior_MCI)
      ratio_NCI <- c(ratio_NCI,current_NCI/bin_size/prior_NCI)
  }
  df_percentage <- data.frame(gene1 = lcpm_gene1,NCI=ratio_NCI,MCI=ratio_MCI,AD=ratio_AD)
  p <- ggplot(data=df_percentage)+xlab("MT-ND1(lcpm)")+ylab("relative ratio")+
    geom_line(aes_string(x="gene1",y="NCI"),col="green")+
     geom_line(aes_string(x="gene1",y="MCI"),col="blue")+
     geom_line(aes_string(x="gene1",y="AD"),col="red")
  ggsave(p,filename = "Fig1/AD percentage change with MT-ND1.png",width = 6, height =2)
  
  
 
  
  
ggplot(data=df_rest,aes_string(x="AOD",y="gene1",col="AD"))+
  geom_point(size=4)+ylab("MT-ND1")+
  scale_colour_manual(values = c("green", "blue", "red"))

ggplot(data=df_rest,aes_string(x="AD",y="gene1",col="AD"))+
  geom_violin()+ylab("MT-ND1")+
  scale_colour_manual(values = c("green", "blue", "red"))

ggplot(data=df_rest)+
  geom_freqpoly(aes(x = gene1,
     y = ..density.., colour = AD))
}
```


## plot MT-ND1 with cell marker genes
```{r}
#Neuron markers
#RBFOX3, NeuN, nueron nuclear marker
#ELAVL4, HuD, neuron specific gene splicing
#ENO2, NSE, cell shape marker
#MAP2,  cell shape marker
#DLGAP1, synapse marker, synaptic activity
#DLG4, synaptic scaffold
#KCNH1, 
#GRIA1~4, AMPA receptor
#GRIN1

#REST
#PRNP
#MIR124-2HG
#TPPP p25alpha, oligodendrocyte marker
#save(df_rest,file="~/Dropbox/human/ROSMAP/RNASeq/df_rest_ROSMAP.Rdata")
gene_current <- "MFN2"  #"TUBA4A"
gene_current_ENSG <- x_ROSMAP$genes[x_ROSMAP$genes$HGNC_SYMBOL == gene_current,"ENSEMBL"][1]
df_rest$gene4 <- lcpm_WithoutFactors[gene_current_ENSG,df_rest$sample]
#df_rest$gene5 <- lcpm_WithoutFactors[gene_current_ENSG,df_rest$sample]

##Plot MT-ND1 and NDUFA1
df_rest_filter <- df_rest  #[df_rest$AD == "NCI",]
lm_fit <- lm(gene5 ~ gene4, data=df_rest_filter)
p <- ggplotRegression(lm_fit) +
                 xlab(gene_current) +
                ylab("MT-ND1")
p
ggplot(data=df_rest,aes_string(x="gene4",y="gene5",col="AD")) +
  geom_point()+theme(legend.position = "right")+xlab(gene_current)+scale_colour_manual(values = c("green", "blue", "red"))
ggsave(p,filename = paste0("Fig1/ND1 and ",gene_current," correlations_ROSMAP.png"),width = 5, height =3)

p
p_density <- ggplot(df_rest,aes(x=gene4,col=AD))  +scale_colour_manual(values = c("green", "blue", "red"))+
   xlab(gene_current) + geom_line(stat="density")+ggtitle("MT-ND1 expression density")+
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position = c(0.8, 0.8),plot.title = element_text(hjust = 0.5),legend.title=element_blank())
p_density
ggsave(p_density,filename = paste0("Fig1/",gene_current," densityMap_Braak_ROSMAP.png"),width = 5, height =3)


#Test REST and APP
if(FALSE){
  df_rest_filter <- df_rest#[df_rest$AD == "NCI",]
lm_fit <- lm(gene5 ~ gene4, data=df_rest_filter)
p <- ggplotRegression(lm_fit) +
                 xlab(gene_current) +
                ylab("APP")
p

ggplot(data=df_rest,aes_string(x="AOD",y="gene4",col="AD")) +
  geom_point()+theme(legend.position = "right")+ylab(gene_current)

}
```

## calculate all gene correlations to MT-ND1
```{r}
df_correlations <- data.frame(gene=c(),cor=c(),rsq=c(),pval=c())
lcpm_ND1 <- lcpm_WithoutFactors["ENSG00000198888",]
for(i in 1:length(lcpm_WithoutFactors[,1])){
  lcpm_current <- lcpm_WithoutFactors[i,]
  lm_current <- lm(lcpm_ND1 ~ lcpm_current)
  gene_current <- rownames(lcpm_WithoutFactors)[i]
  cor_current <- signif(lm_current$coef[2],digits = 5)
  rsq_current <- signif(summary(lm_current)$adj.r.squared,digits = 5)
  p_current <- signif(summary(lm_current)$coef[2,4],digits = 5)
  df_current <- data.frame(gene=gene_current,cor=cor_current,rsq=rsq_current,pval=p_current)
  df_correlations <- rbind(df_correlations, df_current)
}
df_correlations[order(df_correlations$rsq,decreasing = T),] -> df_correlations2
rownames(df_correlations2) <- c(1:length(df_correlations2$gene))
df_correlations2
df_correlations2$gene_symbol <- x_ROSMAP$genes[match(df_correlations2$gene,x_ROSMAP$genes$ENSEMBL),]$HGNC_SYMBOL

df_correlations2[,c("gene_symbol","gene", "cor", "rsq", "pval")] -> df_correlations2

write.csv(df_correlations2,file=paste0("Fig1/correlations_ND1_ROSMAP.csv"))
df_correlations2$pval[800]
{
plot(density(log(df_correlations2$pval)))
abline(v=log(df_correlations2$pval[800]),col="red")
}
```

##Heatmap of MT_genes
```{r}
library(scales)
library(ComplexHeatmap)
library(circlize)

genes_heatmap <- c(MT_human,synapseMarkers)
name_heatmap <- "MT_synapse" #"MT"
MT_genes_ENSG <- x_ROSMAP$genes$ENSEMBL[match(genes_heatmap,x_ROSMAP$genes$HGNC_SYMBOL)]
lcpm_WithoutFactors[MT_genes_ENSG,] -> MT_gene_expr
rownames(MT_gene_expr) <- genes_heatmap

##sort samples by eigen gene value of all MT genes
complex_colors_MTgenes <- rep("grey",length(rownames(MT_gene_expr)))
MEList_MTgenes <- moduleEigengenes((as.matrix(t(MT_gene_expr))), colors = complex_colors_MTgenes)
sample_order_MTgenes <- order(MEList_MTgenes$eigengenes$MEgrey)
names_sample <- rownames(MEList_MTgenes$eigengenes)
##Scaling
  base_mean_y = rowMeans(MT_gene_expr,na.rm = T)  #()
  mat_scaled_y = t(apply(MT_gene_expr, 1, scale))
  max_heat = max(mat_scaled_y)
  min_heat = min(mat_scaled_y)
  rownames(mat_scaled_y)->genes_heatmap
  
##set heatmap top labels
  sample_orders <- sample_order_MTgenes
  #type_apoe = s2c4_ROSMAP_filter2[names_order[sample_orders],"apoe_genotype"]
  #ha_apoe = HeatmapAnnotation(df = data.frame(type = type_apoe),col=list(type = c("APOE34" =  "red", "APOE33" = "grey","APOE23" = "green")) )

  #type_sex = s2c4_ROSMAP_filter2[names_order[sample_orders],"msex"]
  #ha_sex = HeatmapAnnotation(df = data.frame(type = type_sex),col=list(type = c("M" =  "green", "F" = "red")) )
  x_ROSMAP$samples
  type_AD = x_ROSMAP$samples[names_sample[sample_orders],"group"]
  ha_AD = HeatmapAnnotation(df = data.frame(type = type_AD),col=list(type = c("NCI" =  "green", "MCI"="blue","AD" = "red")) )

  ha_y <- HeatmapAnnotation(df = data.frame(Diagnosis = type_AD),
                  col=list(Diagnosis = c("AD" =  "purple", "MCI" = "orange","NCI" = "cyan")
                           ) )

  heatmap_all <- Heatmap( mat_scaled_y[,sample_orders], name = "expression", km = 1, 
                          col = colorRamp2(c(min_heat*0.7, 0, max_heat*0.7), c(muted("blue"), "white", muted("red"))),
                          cluster_columns = F,cluster_rows=F,
                          top_annotation = ha_y, top_annotation_height = unit(4, "mm"), 
                          show_row_names = FALSE, show_column_names = FALSE) +
                          Heatmap(base_mean_y, name = "base_mean", show_row_names = T, row_names_gp = gpar(fontsize = 10), width = unit(5, "mm"),
                                  col = colorRamp2(c(min(base_mean_y), max(base_mean_y)), c("green", "red")) )  #+
  
  #Save heatmap
  {
      pdf(paste0("Fig1/heatmap_",name_heatmap,"_orderedByEigenGenes_ROSMAP.pdf"), height=length(genes_heatmap)*0.25+6, width=12)
      print(heatmap_all)
      dev.off()
   }
```



##generate MTlabel based on the expression level of MT-ND1
```{r}
if(FALSE){
MT_genes <- c(complex1_human,complex2_human,complex3_human,complex4_human,complex5_human)   ##dplyr::select(contains("MT-"))    #startsWith("MT-")
ETC_genes <- c(MT_genes[startsWith(MT_genes,"MT-")],MT_genes[!startsWith(MT_genes,"MT-")])
MT_genes <- MT_genes[startsWith(MT_genes,"MT-")]
MT_genes_ENSG <- dplyr::filter(x_ROSMAP$genes,x_ROSMAP$genes$HGNC_SYMBOL %in% MT_genes)
MT_genes_ENSG <- MT_genes_ENSG$ENSEMBL
}
MT_genes_regulator <- c(MT_human,"TFAM","TFB1M","TFB2M","NRF1","NRF2")

dplyr::filter(x_ROSMAP$genes,x_ROSMAP$genes$HGNC_SYMBOL=="MT-ND1")

expr_currentGene <- lcpm_WithoutFactors["ENSG00000198888",]  #ENSG00000198888, MT-ND1
{
 plot(density(t(expr_currentGene)),main="MT-ND1",xlab("lcpm"))
 abline(v=mean(t(expr_currentGene))+1.0*sd(t(expr_currentGene)),col="red")
 abline(v=mean(t(expr_currentGene))-1.0*sd(t(expr_currentGene)),col="red")
}

{
  p <- ggplot(df_rest,aes(x=gene1))  + #scale_colour_manual(values = c("red","blue", "green"))+
    xlim(18,22)+xlab("lcpm (MT-ND1)") + geom_line(stat="density")+ggtitle("MT-ND1 expression density")+
    geom_vline(xintercept = mean(df_rest$gene1)+1.0*sd(df_rest$gene1), col="red") +
    geom_vline(xintercept = mean(df_rest$gene1)-1.0*sd(df_rest$gene1), col="red")
    #geom_vline(xintercept = mean(gene1)) #theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),legend.position = c(0.8, 0.8),plot.title = element_text(hjust = 0.5),legend.title=element_blank())
  ggsave(p,filename = "Fig1/MT-ND1 density map MThighVslow.png",width = 4, height =3)
}

sum(expr_currentGene > mean(t(expr_currentGene))+1.0*sd(t(expr_currentGene)))  #63
sum(expr_currentGene < mean(t(expr_currentGene)) - 1.0*sd(t(expr_currentGene)))  #76
x_ROSMAP$samples$MTlabel="middle"
tip_high <- rownames(x_ROSMAP$samples) %in% names(expr_currentGene)[expr_currentGene > mean(t(expr_currentGene))+1.0*sd(t(expr_currentGene))]
tip_low <- rownames(x_ROSMAP$samples) %in% names(expr_currentGene)[expr_currentGene < mean(t(expr_currentGene)) - 1.0*sd(t(expr_currentGene))]

x_ROSMAP$samples$MTlabel[tip_high] <- "high"
x_ROSMAP$samples$MTlabel[tip_low] <- "low"

dplyr::filter(x_ROSMAP$genes,x_ROSMAP$genes$HGNC_SYMBOL %in% MT_genes)
x_ROSMAP$samples$MTlabel
```

#DE analysis based on MT labels
```{r}
dim(x_ROSMAP)
y=x_ROSMAP
y$samples$Batch <- relevel(y$samples$Batch,"0")
y$samples$apoe_genotype <- relevel(y$samples$apoe_genotype, "APOE33")
y$samples$msex <- relevel(y$samples$msex,"M")
y$samples$group <- relevel(y$samples$group,"NCI")
y$samples$MTlabel <- factor(y$samples$MTlabel,levels =c("low","middle","high"))
y$samples$groupCom <- as.factor(y$samples$groupCom)
y$samples$groupCom <- relevel(y$samples$groupCom,"NCI.M.APOE33")
#design <- model.matrix(~y$samples$Batch+y$samples$pmi+y$samples$RIN+y$samples$AOD2+y$samples$MTlabel) #+y$samples$groupCom
y$samples$MTlabel2 <- "middle"
y$samples$MTlabel2[y$counts["ENSG00000198888",] <= 404274] <- "low"
y$samples$MTlabel2[y$counts["ENSG00000198888",] >= 692725] <- "high"

y$samples$MTlabel3 <- "high"
y$samples$MTlabel3[y$counts["ENSG00000198888",] <= 533236] <- "low"
#y$samples$MTlabel3[y$counts["ENSG00000198888",] >= 692725] <- "high"

y$counts["ENSG00000198888",]
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#  29592  404274  533236  582620  692725 1863163 
design <- model.matrix(~y$samples$Batch+y$samples$pmi+y$samples$RIN+y$samples$AOD2+y$samples$MTlabel3:y$samples$group) #+y$samples$groupCom
colnames(design) <- gsub("y\\$samples\\$", "", colnames(design))
head(design)
colnames(design) <- gsub("apoe_genotype", "", colnames(design))
head(design)
colnames(design) <- gsub("group", "", colnames(design))
head(design)
colnames(design) <- gsub("Com", "", colnames(design))
head(design)
colnames(design) <- gsub("MTlabel", "MT", colnames(design))
head(design)
colnames(design) <- gsub(":", "_", colnames(design))

rownames(design) <- rownames(y$samples)
if(colnames(design)[1] == "(Intercept)"){
  colnames(design)[1] <- "Intercept"
}
head(design)

```
##Contrast
```{r}
colnames(design)
diagnosis <- c("NCI","MCI","AD")
sex <- c("F","M")
apoe_genotype <- c("APOE33","APOE34","APOE23")

if(FALSE){
  contr.matrix<- makeContrasts(
   ADvsNCI_F_APOE33 = AD.F.APOE33-NCI.F.APOE33, 
   MCIvsNCI_F_APOE33 = MCI.F.APOE33 - NCI.F.APOE33, 
   ADvsMCI_F_APOE33 = AD.F.APOE33 - MCI.F.APOE33, 
   
   ADvsNCI_F_APOE34 = AD.F.APOE34-NCI.F.APOE34, 
   MCIvsNCI_F_APOE34 = MCI.F.APOE34 - NCI.F.APOE34, 
   ADvsMCI_F_APOE34 = AD.F.APOE34 - MCI.F.APOE34,
   
   ADvsNCI_F_APOE23 = AD.F.APOE23-NCI.F.APOE23, 
   MCIvsNCI_F_APOE23 = MCI.F.APOE23 - NCI.F.APOE23, 
   ADvsMCI_F_APOE23 = AD.F.APOE23 - MCI.F.APOE23,
   
   ADvsNCI_M_APOE33 = AD.M.APOE33,   #-NCI.M.APOE33
   MCIvsNCI_M_APOE33 = MCI.M.APOE33, #-NCI.M.APOE33
   ADvsMCI_M_APOE33 = AD.M.APOE33 - MCI.M.APOE33, 
   
   ADvsNCI_M_APOE34 = AD.M.APOE34-NCI.M.APOE34, 
   MCIvsNCI_M_APOE34 = MCI.M.APOE34 - NCI.M.APOE34, 
   ADvsMCI_M_APOE34 = AD.M.APOE34 - MCI.M.APOE34,
   
   ADvsNCI_M_APOE23 = AD.M.APOE23-NCI.M.APOE23, 
   MCIvsNCI_M_APOE23 = MCI.M.APOE23 - NCI.M.APOE23, 
   ADvsMCI_M_APOE23 = AD.M.APOE23 - MCI.M.APOE23,
   
   
   levels = colnames(design))
}

contr.matrix<- makeContrasts(
   highvslow_MT = MThigh, 
   levels = colnames(design))

contr.matrix<- makeContrasts(
   ADvsNCI = AD, 
   levels = colnames(design))
#AD MCI MTmiddle:AD MThigh:AD

contr.matrix<- makeContrasts(
   ADvsNCI_L = MT2low_AD-MT2low_NCI, #
   ADvsNCI_M = MT2middle_AD-MT2middle_NCI,
   ADvsNCI_H = MT2high_AD-MT2high_NCI,
   levels = colnames(design))

contr.matrix<- makeContrasts(
   ADvsNCI_L = MT3low_AD-MT3low_NCI, #
   ADvsNCI_H = MT3high_AD-MT3high_NCI,
   levels = colnames(design))

contr.matrix

c(sum(design[,"MT2low_AD"]),sum(design[,"MT2low_NCI"])) #40,63
c(sum(design[,"MT2middle_AD"]),sum(design[,"MT2middle_NCI"]))#126,89
c(sum(design[,"MT2high_AD"]),sum(design[,"MT2high_NCI"])) # 68,41

c(sum(design[,"MT3low_AD"]),sum(design[,"MT3low_NCI"])) #97 115
c(sum(design[,"MT3high_AD"]),sum(design[,"MT3high_NCI"])) # 137  78
```

```{r,eval=TRUE}
par(mfrow=c(1,2))
v <- voom(y, design, plot=TRUE)
dim(v)

```


```{r,eval=TRUE}
vfit <- lmFit(v, design)
vfit_contr <- contrasts.fit(vfit, contrasts=contr.matrix)
#vfit_contr_APOE <- contrasts.fit(vfit, contrasts=contr.matrix_APOE)
#vfit_contr_Gender <- contrasts.fit(vfit, contrasts=contr.matrix_Gender)

efit <- eBayes(vfit_contr)
#efit_APOE <- eBayes(vfit_contr_APOE)
#efit_Gender <- eBayes(vfit_contr_Gender)
```

```{r}
plotSA(efit, main="Final model: Mean−variance trend")

```

```{r}
summary(decideTests(efit))

```
```{r}
head(coef(efit))
topTable(efit,number=15,coef="highvslow_MT",adjust.method = "BH")[,c(1,3,5:11)]
topTable(efit,number=15,coef="ADvsNCI",adjust.method = "BH",confint=TRUE)[,c(1,3,7:13)]

```

#Save DEG the results
```{r}
#summary(decideTests(efit))
topTable(efit,coef=1,number=Inf,sort.by="B",p.value=0.01, lfc=log2(1.1)) [1:5,] #coef=1
comparisons <- colnames(summary(decideTests(efit)))

for(i in 1:length(comparisons)){
  current_compare <- comparisons[i]
  currentDEGs <- topTable(efit,number=Inf,coef=current_compare,adjust.method = "BH",confint=TRUE)
  write.csv(currentDEGs,file=paste0("Fig1/",current_compare,"_AD_mitoLevels2.csv"))
}

```

#GO analysis
```{r}
DElist <- topTable(efit,coef=1,number=Inf,sort.by="B",p.value=0.01, lfc=log2(1.8)) #coef=1
DElist[DElist$logFC>0,"ENSEMBL"] -> up_regulated_MT  #439
length(up_regulated_MT)
#up_regulated_MT[!up_regulated_MT %in% MT_genes_ENSG] -> up_regulated_MT
DElist[DElist$logFC<0,"ENSEMBL"] -> down_regulated_MT #410
length(down_regulated_MT)

geneSets <- list(up=up_regulated_MT,down=down_regulated_MT)
GO_types <- c("BP","MF","CC")    
i=2
for(i in 1:2){
current_set <- geneSets[[i]]
which.color <- names(geneSets[i])
keytypes(org.Hs.eg.db)
trait_allGenes_entrez <- mapIds(org.Hs.eg.db,
                     keys=current_set,
                     column="ENTREZID", #ENSEMBL
                     keytype="ENSEMBL", #SYMBOL
                     multiVals="first")
     trait_allGenes_entrez[!is.na(trait_allGenes_entrez)] -> trait_allGenes_entrez
     
     j=3
     for(j in 1:3){
       GO_type = GO_types[j]
       ego_allgene <- enrichGO(gene          = trait_allGenes_entrez,
                # universe      = names(geneList),
                OrgDb         = org.Hs.eg.db,
                ont           = GO_type,  #BP, MF (Molecular Function), CC (Cellular Components) 
                pAdjustMethod = "BH",
                pvalueCutoff  = 0.1,
                qvalueCutoff  = 0.1,
                readable      = TRUE)
       length_positive <- length(as.data.frame(ego_allgene)[,1])
       if(length_positive >=1){
          GO_BP <- as.data.frame(ego_allgene)
          write.xlsx(GO_BP,paste0("Fig1/",which.color,"_AllgeneGO",length(trait_allGenes_entrez),"_genes_GO_",GO_type,"_ROSMAP_MT.xlsx"))
          cnetplot <- cnetplot(ego_allgene,showCategory = 5)
          ggsave(filename=paste0("Fig1/",which.color,"_cnetplot_",GO_type,"_ROSMAP_MT.png"),cnetplot,width=8,height=4,dpi=300)
          dot_plot <- dotplot(ego_allgene,showCategory = 10)
          ggsave(filename=paste0("Fig1/",which.color,"_dotPlot_",GO_type,"_ROSMAP_MT_top10_2.png"),dot_plot,width=7,height=5,dpi=300)
          dot_plot2 <- dotplot(ego_allgene,showCategory = min(30,length_positive))
          ggsave(filename=paste0("Fig1/",which.color,"_dotPlot_",GO_type,"_ROSMAP_MT.png"),dot_plot2,width=10,height=ceiling(min(30,length_positive)/2),dpi=300)
       }
        cat(paste0(which.color,"\t",GO_type,"\t","All","\n"))
     }
}


#For up_MT genes, remove Mito part
up_regulated_MT[!up_regulated_MT %in% MT_genes_ENSG] -> up_regulated_MT2
{
  current_set <- up_regulated_MT2
which.color <- "up_regulated_MT_removeMT"
keytypes(org.Hs.eg.db)
trait_allGenes_entrez <- mapIds(org.Hs.eg.db,
                     keys=current_set,
                     column="ENTREZID", #ENSEMBL
                     keytype="ENSEMBL", #SYMBOL
                     multiVals="first")
     trait_allGenes_entrez[!is.na(trait_allGenes_entrez)] -> trait_allGenes_entrez
     
     j=3
     for(j in 1:3){
       GO_type = GO_types[j]
       ego_allgene <- enrichGO(gene          = trait_allGenes_entrez,
                # universe      = names(geneList),
                OrgDb         = org.Hs.eg.db,
                ont           = GO_type,  #BP, MF (Molecular Function), CC (Cellular Components) 
                pAdjustMethod = "BH",
                pvalueCutoff  = 0.1,
                qvalueCutoff  = 0.4,
                readable      = TRUE)
       length_positive <- length(as.data.frame(ego_allgene)[,1])
       if(length_positive >=1){
          GO_BP <- as.data.frame(ego_allgene)
          write.xlsx(GO_BP,paste0("Fig1/",which.color,"_AllgeneGO",length(trait_allGenes_entrez),"_genes_GO_",GO_type,"_ROSMAP_MT_removeMT.xlsx"))
          cnetplot <- cnetplot(ego_allgene,showCategory = 5)
          ggsave(filename=paste0("Fig1/",which.color,"_cnetplot_",GO_type,"_ROSMAP_MT_removeMT.png"),cnetplot,width=8,height=4,dpi=300)
          dot_plot <- dotplot(ego_allgene,showCategory = 10)
          ggsave(filename=paste0("Fig1/",which.color,"_dotPlot_",GO_type,"_ROSMAP_MT_top10_removeMT.png"),dot_plot,width=10,height=5,dpi=300)
          dot_plot2 <- dotplot(ego_allgene,showCategory = min(30,length_positive))
          ggsave(filename=paste0("Fig1/",which.color,"_dotPlot_",GO_type,"_ROSMAP_MT_removeMT.png"),dot_plot2,width=10,height=ceiling(min(30,length_positive)/2),dpi=300)
       }
        cat(paste0(which.color,"\t",GO_type,"\t","All","\n"))
     }
}
```


#Camera analysis

```{r}
##Read GMT
#c5.all.v7.0.entrez.gmt  #GO
#c2.all.v7.0.entrez.gmt  #KEGG
#c1.all.v7.0.entrez.gmt  #position
#c3.all.v7.0.entrez      #motif genesets
current_set <- "c3.all.v7.0.entrez"
gmt_file <- paste0("~/Dropbox/RaymondTools/GSEA/gmt/",current_set,".gmt")
gmt <- read_lines(gmt_file)
str(gmt)
n_gmt <-0
current_gmt_total <- list()
for(currentLine in gmt){
  tmp_gmt <- strsplit(currentLine,split="\t")
  tmp_gmt <- tmp_gmt[[1]]
  n_gmt <- n_gmt +1
  current_gmt_total <- c(current_gmt_total, list(tmp_gmt[3:length(tmp_gmt)])  )  #list.append(Hs.C5.v7.0,d=tmp_gmt[3:length(tmp_gmt)])
  names(current_gmt_total)[length(current_gmt_total)] <- tmp_gmt[1]
}
Hs.C3.v7.0 <- current_gmt_total
save(Hs.C3.v7.0,file="~/Dropbox/RaymondTools/GSEA/human_c3.all.v7.0.rdata")



#Hs.C2.v7.0
#Hs.C5.v7.0
#Hs.C1.v7.0
Hs.C1.v7.0[[1]]
Hs.C3.v7.0
```

```{r}
library("xlsx")

load("~/Dropbox/RaymondTools/GSEA/human_c5.all.v7.0.rdata") 
idx_c5 <- ids2indices(Hs.C5.v7.0,id=v$genes$ENTREZID) 
load("~/Dropbox/RaymondTools/GSEA/human_c2.all.v7.0.rdata") 
idx_c2 <- ids2indices(Hs.C2.v7.0,id=v$genes$ENTREZID) 
idx_c2_KEGG <- idx_c2[startsWith(names(idx_c2),"KEGG_")]
load("~/Dropbox/RaymondTools/GSEA/human_c1.all.v7.0.rdata") 
idx_c1 <- ids2indices(Hs.C1.v7.0,id=v$genes$ENTREZID) 
#GO_MYELIN_SHEATH
load("~/Dropbox/RaymondTools/GSEA/human_c3.all.v7.0.rdata") 
idx_c3 <- ids2indices(Hs.C3.v7.0,id=v$genes$ENTREZID) 

cam.MThigh_GO <- camera(v,idx_c5,design,contrast=contr.matrix[,1]) 
head(cam.MThigh_GO,100)
cam.MThigh_GO_down <- cam.MThigh_GO[cam.MThigh_GO$Direction=="Down",]
cam.MThigh_GO_up <- cam.MThigh_GO[cam.MThigh_GO$Direction=="Up",]
#Save camera result
write.xlsx2(cam.MThigh_GO_down, file=paste0("Fig1/","GO_","camera.xlsx"), 
                 sheetName="GO_down", row.names=TRUE)#current_compare
write.xlsx2(cam.MThigh_GO_up, file=paste0("Fig1/","GO_","camera.xlsx"), 
                 sheetName="GO_up", append=TRUE, row.names=TRUE)#current_compare



cam.MThigh_KEGG <- camera(v,idx_c2_KEGG,design,contrast=contr.matrix[,1]) 
head(cam.MThigh_KEGG,100)
cam.MThigh_KEGG_down <- cam.MThigh_KEGG[cam.MThigh_KEGG$Direction=="Down",]
cam.MThigh_KEGG_up <- cam.MThigh_KEGG[cam.MThigh_KEGG$Direction=="Up",]
#Save camera result
write.xlsx2(cam.MThigh_KEGG_down, file=paste0("Fig1/","KEGG_","camera.xlsx"), 
                 sheetName="KEGG_down", row.names=TRUE)#current_compare
write.xlsx2(cam.MThigh_KEGG_up, file=paste0("Fig1/","KEGG_","camera.xlsx"), 
                 sheetName="KEGG_up", append=TRUE, row.names=TRUE)#current_compare


camKEGG.MThigh_GenomePosition <- camera(v,idx_c1,design,contrast=contr.matrix[,1]) 
head(camKEGG.MThigh_GenomePosition,100)
camKEGG.MThigh_GenomePosition_down <- camKEGG.MThigh_GenomePosition[camKEGG.MThigh_GenomePosition$Direction=="Down",]
camKEGG.MThigh_GenomePosition_up <- camKEGG.MThigh_GenomePosition[camKEGG.MThigh_GenomePosition$Direction=="Up",]
#Save camera result
write.xlsx2(camKEGG.MThigh_GenomePosition_down, file=paste0("Fig1/","GenomePosition_","camera.xlsx"), 
                 sheetName="KEGG_down", row.names=TRUE)#current_compare
write.xlsx2(camKEGG.MThigh_GenomePosition_up, file=paste0("Fig1/","GenomePosition_","camera.xlsx"), 
                 sheetName="KEGG_up", append=TRUE, row.names=TRUE)#current_compare


cam.MThigh_motif <- camera(v,idx_c3,design,contrast=contr.matrix[,1]) 
head(cam.MThigh_motif,100)
cam.MThigh_motif_down <- cam.MThigh_motif[cam.MThigh_motif$Direction=="Down",]
cam.MThigh_motif_up <- cam.MThigh_motif[cam.MThigh_motif$Direction=="Up",]
#Save camera result
write.xlsx2(cam.MThigh_motif_down, file=paste0("Fig1/","Motif","camera.xlsx"), 
                 sheetName="Motif_down", row.names=TRUE)#current_compare
write.xlsx2(cam.MThigh_motif_up, file=paste0("Fig1/","Motif","camera.xlsx"), 
                 sheetName="Motif_up", append=TRUE, row.names=TRUE)#current_compare

```



#Convert ENSG to gene symbols
```{r}

expr_ROSMAP <- data.frame(lcpm_WithoutFactors)
ROSMAP_ids <- colnames(x_ROSMAP)
{
s2c4_ROSMAP <- s2c4_ROSMAP_yshang
s2c4_ROSMAP$rnaseq_id <- paste0("X",s2c4_ROSMAP$rnaseq_id )
ROSMAP_ids[(ROSMAP_ids %in% s2c4_ROSMAP$rnaseq_id )] -> ROSMAP_ids_filter
ROSMAP_ids_filter[ROSMAP_ids_filter %in% colnames(expr_ROSMAP)] -> ROSMAP_ids_filter
s2c4_ROSMAP <- s2c4_ROSMAP[s2c4_ROSMAP$rnaseq_id %in% ROSMAP_ids_filter,]
# [1] "X01_120405"   "X150_120419"  "X164_120423"  "X20_120411"   "X218_120425"  "X266_120430"  "X276_120430"  "X411_120503"  "X448_120507"  "X451_120507"  "X517_120515" 
#[12] "X665_120530"  "X674_120530"  "X700_120605"  "X492_120515"  "R24_131017"   "redo4_140501" "X892_130923" 
expr_ROSMAP[,ROSMAP_ids_filter] ->expr_ROSMAP_filter
head(expr_ROSMAP_filter)[1:5,1:5]
ENSGs <- rownames(expr_ROSMAP_filter)


ENSGs_symbol <- mapIds(org.Hs.eg.db,
                     keys=ENSGs,
                     column="SYMBOL", #ENSEMBL
                     keytype="ENSEMBL", #SYMBOL
                     multiVals="first")

expr_ROSMAP_filter[ENSGs[!is.na(ENSGs_symbol)],] -> expr_ROSMAP_filter2

head(expr_ROSMAP_filter2)
##  'select()' returned 1:many mapping between keys and columns
ENSGs[!is.na(ENSGs_symbol)] -> ENSGs2
ENSGs_symbol2 <- mapIds(org.Hs.eg.db,
                     keys=ENSGs2,
                     column="SYMBOL", #ENSEMBL
                     keytype="ENSEMBL", #SYMBOL
                     multiVals="first")
ENSGs2[!duplicated(ENSGs_symbol2)] -> ENSGs2
expr_ROSMAP_filter2[ENSGs2,] -> expr_ROSMAP_filter2
ENSGs_symbol2 <- mapIds(org.Hs.eg.db,
                     keys=ENSGs2,
                     column="SYMBOL", #ENSEMBL
                     keytype="ENSEMBL", #SYMBOL
                     multiVals="first")
rownames(expr_ROSMAP_filter2) <- ENSGs_symbol2

head(expr_ROSMAP_filter2)[1:5,1:5]

colnames(expr_ROSMAP_filter2)[1:4]
length(s2c4_ROSMAP$rnaseq_id) - length(colnames(expr_ROSMAP_filter2))
unique(s2c4_ROSMAP$msex)  #"F" "M"
unique(s2c4_ROSMAP$apoe_genotype) #"APOE24" "APOE33" "APOE23" "APOE34" "APOE22" "APOE44"
unique(s2c4_ROSMAP$cogdx)      # 4 1 2 5 3
unique(s2c4_ROSMAP$AD)         #"AD"  "NCI" "MCI"

s2c4_ROSMAP_filter <- s2c4_ROSMAP[s2c4_ROSMAP$apoe_genotype %in% c("APOE33","APOE23","APOE34"),]
s2c4_ROSMAP_filter$condition <- paste(s2c4_ROSMAP_filter$AD,s2c4_ROSMAP_filter$msex,s2c4_ROSMAP_filter$apoe_genotype,sep=".")
unique(s2c4_ROSMAP_filter$condition)
condition_order <- c("NCI.M.APOE23","NCI.M.APOE33","NCI.M.APOE34",
                     "NCI.F.APOE23","NCI.F.APOE33","NCI.F.APOE34",
                     "MCI.M.APOE23","MCI.M.APOE33","MCI.M.APOE34",
                     "MCI.F.APOE23","MCI.F.APOE33","MCI.F.APOE34",
                     "AD.M.APOE23", "AD.M.APOE33", "AD.M.APOE34",
                     "AD.F.APOE23", "AD.F.APOE33", "AD.F.APOE34"
                      )
s2c4_ROSMAP_filter$condition <- factor(s2c4_ROSMAP_filter$condition, levels =condition_order)
s2c4_ROSMAP_filter$rnaseq_id[order(s2c4_ROSMAP_filter$condition)] -> sample_ordered

expr_ROSMAP_filter2[,sample_ordered] -> expr_ROSMAP_filter3
head(expr_ROSMAP_filter3)[1:5,1:5]
length(colnames(expr_ROSMAP_filter2))
length(colnames(expr_ROSMAP_filter3))
length(s2c4_ROSMAP_filter$rnaseq_id)
lapply(expr_ROSMAP_filter3,as.numeric,2) -> expr_ROSMAP_filter4
data.frame(expr_ROSMAP_filter4,check.names = F) -> expr_ROSMAP_filter4
rownames(expr_ROSMAP_filter4) <- rownames(expr_ROSMAP_filter3)
#s2c4_ROSMAP_filter
#expr_ROSMAP_filter4
str(expr_ROSMAP_filter4[1:5,1:5])
}
```

##convert unmapped symbols
```{r}

test_symbol <- complex5_human
test_ENTREZID <- mapIds(org.Hs.eg.db,
                     keys=test_symbol,
                     column="ENTREZID", #ENSEMBL,ENTREZID
                     keytype="SYMBOL", #SYMBOL
                     multiVals="first")

test_symbol[is.na(test_ENTREZID)]
outliers_human_old <- c("MT-ND1","MT-ND2","MT-ND3", "MT-ND4","MT-ND4L", "MT-ND5","MT-ND6","MT-CYB","MT-CO1", "MT-CO2", "MT-CO3","MT-ATP6", "MT-ATP8")
outliers_human_new <- c("ND1","ND2" ,"ND3","ND4","ND4L","ND5","ND6","CYTB","COX1","COX2","COX3","ATP6","ATP8")

 mapIds(org.Hs.eg.db,
                     keys=outliers_human_new,
                     column="ENTREZID", #ENSEMBL,ENTREZID
                     keytype="SYMBOL", #SYMBOL
                     multiVals="first")
```




```{r}
allgenes_detected <- rownames(expr_ROSMAP_filter4)
for(i in 1:length(outliers_human_new)){
  allgenes_detected <- gsub(outliers_human_new[i],outliers_human_old[i],allgenes_detected)
}

allgenes_detected <- gsub("MT-MT-ND4L","MT-ND4L",allgenes_detected)

rownames(expr_ROSMAP_filter4) <- allgenes_detected
allgenes_detected[startsWith(allgenes_detected,"MT-")]
samples_all <- colnames(expr_ROSMAP_filter4)
s2c4_ROSMAP_filter2 <- s2c4_ROSMAP_filter
rownames(s2c4_ROSMAP_filter2) <- s2c4_ROSMAP_filter2$rnaseq_id
s2c4_ROSMAP_filter2[samples_all,] -> s2c4_ROSMAP_filter2
#s2c4_ROSMAP_filter2 
#expr_ROSMAP_filter4
```

##plot heatmap
```{r}
xpand <- function(d) do.call("expand.grid", rep(list(1:nrow(d)), 2))
euc_norm <- function(x) sqrt(sum(x^2))
euc_dist <- function(mat, weights=1) {
  iter <- xpand(mat)
  vec <- mapply(function(i,j) euc_norm(weights*(mat[i,] - mat[j,])), 
                iter[,1], iter[,2])
  matrix(vec,nrow(mat), nrow(mat))
}
```

```{r}
library(scales)
library(ComplexHeatmap)
library(circlize)

allgenes_detected_ROSMAP <- rownames(expr_ROSMAP_filter4)
sample_order <- s2c4_ROSMAP_filter2$rnaseq_id

genes_detected_ROSMAP <- rownames(expr_ROSMAP_filter4)

MT_genes <- c(complex1_human,complex2_human,complex3_human,complex4_human,complex5_human)   ##dplyr::select(contains("MT-"))    #startsWith("MT-")
ETC_genes <- c(MT_genes[startsWith(MT_genes,"MT-")],MT_genes[!startsWith(MT_genes,"MT-")])
MT_genes <- MT_genes[startsWith(MT_genes,"MT-")]
#MT_genes_regulator <- c(MT_genes,"TFAM","TFB1M","TFB2M","NRF1","NRF2")

MT_genes_regulator <- c(MT_genes,"TFAM","TFB1M","TFB2M","NRF1","NRF2","DLG4","DLGAP1","SYNGAP1","SHANK1","SHANK2","SHANK3")

MT_genes_MTfusion <- c(MT_genes,"DNM1L","MFN2","MFN1","OPA1","DLGAP1","KCNH1","GRIN2A","GRIN2B","LRRTM3","GRIA1","GRIA2","GRIA3","GRIA4")

MT_genes_CellType <- c(MT_genes,"NRGN","GAD1","AQP4","MBP","CSF1R","CD74","VCAN","FLT1","AMBP","DLGAP1","KCNH1","GRIN2A","GRIN2B")


complexes <- list(complex1 =complex1_human, complex2=complex2_human, complex3=complex3_human, 
                  complex4=complex4_human,complex5=complex5_human,complex_MT=MT_genes,
                  complex_ETC=ETC_genes,complex_MT_regulator = MT_genes_regulator,
                  MT_genes_MTfusion=MT_genes_MTfusion,
                  MT_genes_CellType=MT_genes_CellType)


#sort samples by eigen gene value of all MT genes
expr_ROSMAP_filter4[MT_genes,] -> expr_ROSMAP_MTgenes
complex_colors_MTgenes <- rep("grey",length(rownames(expr_ROSMAP_MTgenes)))
MEList_MTgenes <- moduleEigengenes((as.matrix(t(expr_ROSMAP_MTgenes))), colors = complex_colors_MTgenes)
sample_order_MTgenes <- order(MEList_MTgenes$eigengenes$MEgrey)

complex_select <- complex5_human
complex_select_name <- "complex5_human_ROSMAP_Reseq"

complex_select <- ETC_genes
complex_select_name <- "ETCgenes_human_ROSMAP_Reseq"

complex_select <- MT_genes
complex_select_name <- "MTgenes_human_ROSMAP_Reseq"

complex_select <- MT_genes_regulator
complex_select_name <- "MTgenesRegulator_human_ROSMAP_Reseq_CovarRemove"

order_MTgenes = TRUE
row_cluster= FALSE#TRUE
i=1


for(i in 1:length(complexes)){
  complex_select_name=names(complexes[i])
  complex_select <- complexes[[i]]
  row_select <- which(genes_detected_ROSMAP %in% complex_select)
  complex_select_filter <- genes_detected_ROSMAP[row_select]
  complex_select[complex_select %in% complex_select_filter] -> complex_select_filter
  col_select <- which(s2c4_ROSMAP_filter2$AD %in% c("NCI","MCI","AD") )  #& s2c_expr$msex %in% "F"
  mat_y <- expr_ROSMAP_filter4[row_select,col_select]#lcpm_mds[row_select,]  #v$E[row_select,]
  colnames(mat_y) -> names_tmp

  names_order <- names_tmp  #c(c1,c2,c3)
  mat_y[complex_select_filter,names_order] -> mat_y
  
  base_mean_y = rowMeans(mat_y,na.rm = T)  #()
  mat_scaled_y = t(apply(mat_y, 1, scale))
  max_heat = max(mat_scaled_y)
  min_heat = min(mat_scaled_y)
  rownames(mat_scaled_y)->gene_names
  
  #cluster by weighted distance
  if(!order_MTgenes){
    n_MT = sum(startsWith(gene_names,"MT-"))
    n_nuc = length(gene_names)-n_MT
    weight <- c(rep(n_nuc,n_MT),rep(n_MT,n_nuc))#complex4
    if(n_MT==0 | n_nuc==0){
        weight=1
    }
    sampledist_weighted <- euc_dist(mat=t(mat_scaled_y[,]), weights=weight)  #1
    sampledist_weighted
    cluster_samples <- hclust(as.dist(sampledist_weighted),method = "complete") #average  complete
    sample_orders <- cluster_samples$order
  }else{
    sample_orders <- sample_order_MTgenes
  }

  #set heatmap top labels
  type_apoe = s2c4_ROSMAP_filter2[names_order[sample_orders],"apoe_genotype"]
  ha_apoe = HeatmapAnnotation(df = data.frame(type = type_apoe),col=list(type = c("APOE34" =  "red", "APOE33" = "grey","APOE23" = "green")) )

  type_sex = s2c4_ROSMAP_filter2[names_order[sample_orders],"msex"]
  ha_sex = HeatmapAnnotation(df = data.frame(type = type_sex),col=list(type = c("M" =  "green", "F" = "red")) )

  type_AD = s2c4_ROSMAP_filter2[names_order[sample_orders],"AD"]
  ha_AD = HeatmapAnnotation(df = data.frame(type = type_AD),col=list(type = c("NCI" =  "green", "MCI"="white","AD" = "red")) )

  ha_y <- HeatmapAnnotation(df = data.frame(Diagnosis = type_AD,Sex=type_sex,APOE=type_apoe),
                  col=list(Diagnosis = c("AD" =  "purple", "MCI" = "orange","NCI" = "cyan"),
                           Sex = c("F" =  "green", "M" = "blue"),
                           APOE= c("APOE33" =  "grey", "APOE23" = "pink","APOE34" = "magenta")) )

  heatmap_all <- Heatmap( mat_scaled_y[,sample_orders], name = "expression(Scaled)", km = 1, 
                          col = colorRamp2(c(min_heat*0.7, 0, max_heat*0.7), c(muted("blue"), "white", muted("red"))),
                          cluster_columns = !(order_MTgenes),cluster_rows=row_cluster,
                          top_annotation = ha_y, top_annotation_height = unit(4, "mm"), 
                          show_row_names = FALSE, show_column_names = FALSE) +
                          Heatmap(base_mean_y, name = "base_mean", show_row_names = T, row_names_gp = gpar(fontsize = 10), width = unit(5, "mm"),
                                  col = colorRamp2(c(min(base_mean_y), 0, max(base_mean_y)), c("green", "white", "red")) )  #+

  #heatmap_all
  if(order_MTgenes){
      pdf(paste0("~/Dropbox/human/RNAResequence/ROSMAP/heatmap/heatmap_","MT_genes_order_",ifelse(row_cluster,"Row","NoRow"),"_Clustering_",complex_select_name,"yshang.pdf"), height=length(complex_select_filter)*0.25+6, width=15)
      print(heatmap_all)
      dev.off()
  }else{
      pdf(paste0("~/Dropbox/human/RNAResequence/ROSMAP/heatmap/heatmap_","weightedCol_",ifelse(row_cluster,"Row","NoRow"),"_Clustering_",complex_select_name,"yshang.pdf"), height=length(complex_select_filter)*0.25+6, width=15)
      print(heatmap_all)
      dev.off()
  }
}



```



###Gene-trait relationships 
  Row is module，Collumn is trait。correlation coefficient and pvalues are show in the squares
  
```{r}
#RINcontinuous,pmi,age_death,educ,braaksc,cogdx,ceradsc,Batch
x_lab="RINcontinuous"
trail_current <- s2c4_ROSMAP_filter2[names(expr_currentGene),x_lab]
trail_current2 <- trail_current^2
str(t(expr_currentGene))
expr_currentGene2 <- (expr_currentGene)
#expr_currentGene2 <- expr_currentGene2[,1]
str(expr_currentGene2)
str(trail_current)
cor(expr_currentGene2, trail_current)
plot(x=trail_current,y=expr_currentGene2,xlab=x_lab,ylab="MT-ND1")
lm.fit <- lm(expr_currentGene2~trail_current)
summary(lm.fit)

residuals_gene <- lm.fit$residuals
{
par(mfrow=c(1, 2))
plot(x=c(1:588),y=expr_currentGene2[sample_orders],main="Normalized_lcpm")
plot(x=c(1:588),y=residuals_gene[sample_orders],main="Residuals")
}
plot(x=trail_current,y=residuals_gene,xlab=x_lab,ylab="MT-ND1",main="residual")

summary(lm(residuals_gene~trail_current))
```
  
```{r}
trail_selected <- c("RINcontinuous","pmi","age_death","educ","braaksc","cogdx","ceradsc","Batch")
expr_currentSet <- expr_ROSMAP_filter4[MT_genes,]

trail_currentSet <- s2c4_ROSMAP_filter2[names(expr_currentGene),trail_selected]


nSamples=length(trail_currentSet$RINcontinuous)
moduleTraitCor_noFP <- cor( t(expr_currentSet), trail_currentSet, use = "p");
moduleTraitPvalue_noFP = corPvalueStudent(moduleTraitCor_noFP, nSamples); 
textMatrix_noFP <- paste(signif(moduleTraitCor_noFP, 2), "\n(", signif(moduleTraitPvalue_noFP, 1), ")", sep = ""); 
{
pdf(file="MTgene-trait relationships_ROSMAP_CovarRemove.pdf",width=8,height=8)
par(mar = c(4.5, 8.5, 2, 2)); 
labeledHeatmap(Matrix = moduleTraitCor_noFP, 
               xLabels = colnames(trail_currentSet), 
               yLabels = rownames(expr_currentSet), 
               ySymbols = rownames(expr_currentSet), 
               colorLabels = FALSE, 
               colors = blueWhiteRed(50), 
               textMatrix = textMatrix_noFP,
               setStdMargins = FALSE, 
               cex.text = 0.65, 
               zlim = c(-1,1), 
               main = paste("Gene-trait relationships")) 

dev.off()
}
```

##Remove multiple factors effect and plot again


###Plot
```{r}
#RINcontinuous,pmi,age_death,educ,braaksc,cogdx,ceradsc,Batch
expr_currentGene <- lcpm_WithoutFactors["ENSG00000198888",] #MT-ND1
expr_currentGene <- expr_currentGene[names(expr_currentGene) %in% s2c4_ROSMAP_filter2$rnaseq_id]
x_lab="RINcontinuous"
trail_current <- s2c4_ROSMAP_filter2[names(expr_currentGene),x_lab]
trail_current2 <- trail_current^2
str(t(expr_currentGene))
expr_currentGene2 <- (expr_currentGene)
#expr_currentGene2 <- expr_currentGene2[,1]
str(expr_currentGene2)
str(trail_current)
cor(expr_currentGene2, trail_current)
plot(x=trail_current,y=expr_currentGene2,xlab=x_lab,ylab="MT-ND1")
lm.fit <- lm(expr_currentGene2~trail_current)
summary(lm.fit)

residuals_gene <- lm.fit$residuals
{
par(mfrow=c(1, 2))
plot(x=c(1:length(expr_currentGene2)),y=expr_currentGene2[sample_orders],main="Normalized_lcpm")
plot(x=c(1:588),y=residuals_gene[sample_orders],main="Residuals")
}
plot(x=trail_current,y=residuals_gene,xlab=x_lab,ylab="MT-ND1",main="residual")

summary(lm(residuals_gene~trail_current))
```

